The Calculus of Linear Constructions

### 2021-10-20 writing/lics22/lics22.tex
lics22

\lstset{basicstyle=\scriptsize,language=Coq,showstringspaces=false}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

@@

  \begin{figure}[h]
  \vspace{-1.2em}
  \caption{Inductive Type Definition and Elimination}
  \begin{lstlisting}
  (* Sigma Definition *)
  Inductive Sigma (A: Type) (F: A -> Type): Type :=
  | pair : (x: A) -> F x -> Sigma A F.

  (* Tensor Definition *)
  Inductive Tensor (A: Linear) (B: Linear): Linear :=
  | tpair: A -> B -> Tensor A B.

  (* FTensor Definition *)
  Inductive FTensor (A: Type) (F: A -> Linear): Linear :=
  | fpair: (x: A) -> F x -> FTensor A F.

  (* Sigma Elimination
     x and y are unrestricted in clause body *)
  match m with
  | pair x y => (* clause body *)
  end

  (* Tensor Elimination
     x and y must be used once in clause body *)
  match m with
  | tpair x y => (* clause body *)
  end

  (* FTensor Elimination
     x is unrestricted in clause body, 
     y must be used once in clause body *)
  match m with
  | fpair x y => (* clause body *)
  end
  \end{lstlisting}
  \label{inductive}
  \Description{}
  \end{figure}

@@

  Though we have not implemented program extraction for CLC yet, our preliminary experiments hint at the possibility of compiling the state axioms to constant time imperative operations after proof erasure. This can greatly increase the runtime efficiency of CLC over pure functional languages when used for programming. More examples are available in the first author's Github repository.

@@

  Figure \ref{list} presents a CPS append function for linear lists. In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}. Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$. Basically, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.

>>>>

\lstset{basicstyle=\small,language=Coq,showstringspaces=false}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]

@@

  \begin{figure}[h]
  \vspace{-1em}
  \caption{Inductive Type Definition and Elimination}
  \begin{lstlisting}
  (* Sigma Definition *)
  Inductive Sigma (A: U) (F: A -> U): U :=
  | pair : (x: A) -> F x -> Sigma A F.

  (* Tensor Definition *)
  Inductive Tensor (A: L) (B: L): L :=
  | tpair: A -> B -> Tensor A B.

  (* FTensor Definition *)
  Inductive FTensor (A: U) (F: A -> L): L :=
  | fpair: (x: A) -> F x -> FTensor A F.

  (* Sigma Elimination *)
  match m with
  | pair x y => (* clause body *)
  end

  (* Tensor Elimination *)
  match m with
  | tpair x y => (* clause body *)
  end

  (* FTensor Elimination *)
  match m with
  | fpair x y => (* clause body *)
  end
  \end{lstlisting}
  \vspace{-1em}
  \label{inductive}
  \Description{}
  \end{figure}

@@

  Though we have not implemented program extraction for CLC yet, our preliminary experiments show the feasibility of compiling the state axioms to constant time imperative operations after proof erasure. This can greatly improve the runtime efficiency of CLC over pure functional languages when used for programming. More complex examples are available in the first author's Github repository.

@@

  Figure \ref{list} presents a CPS append function for linear lists. In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}. Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$. Basically, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place. Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once. Using linear types for continuations is surprisingly accurate as control-flow itself is linear.
